Immutable Source-Mapped Abstract Syntax Tree:\ \, \\ A Design Pattern for Refactoring Engine APIs

### 2017-09-26 paper.tex
Fixed ACM Reference Format month/year
Added github repo
Variant explicit names

% Metadata Information
\acmVolume{0}
\acmNumber{0}
\acmArticle{0}
\articleSeq{0}
\acmYear{2013}
\acmMonth{5}

@@

    \item \textit{Work with the ASTs in a functional style, where every edit
        returns a new AST.} So, a call to field.removeFromTree() would not
        change the existing AST, but rather it would return a new AST that has
        the given field omitted. When dealing with immutable data structures,
        functional programming may often be the natural solution. However,
        in the case of ASTs, a functional style risks reintroducing many of 
        the problems inherent to mutable trees. File location of nodes may
        be disrupted in the new tree, pointers to nodes in the original tree may
        be rendered 
        irrelevant, semantic information may be invalidated, etc. Those are in
        addition to the difficulties of functional programming, such as memory
        consumption and copying, especially with very large ASTs. This is
        a valid refactoring API design pattern, but any implementation using 
        this variant must carefully consider these problems.

    \item \textit{Provide an API that allows clients to specify what changes
        they want to make in terms of AST nodes.}  This API can then translate these into string edits.  For example, an API might offer a \textit{Rewriter} 
class. The client would inform the \textit{Rewriter} that it wants to 
effectively delete a particular \textit{FieldDeclaration} node from the AST:
\textit{rewriter.delete(field)}.  When it is ready to manipulate the source
code, it delegates this task to the rewriter:
\textit{rewriter.rewrite(sourceCode)}.  This allows multiple edits to be made 
without conflicting with each other. This is illustrated in
Figure~\ref{fig:seq-diagram}.  The Eclipse Java and C/C++ Development
Tools~\cite{cdt-refactoring} take this approach (with slightly different class
and method names).

@@

\section{Example}
We created a plugin for Eclipse that exhibits the \emph{Rewriter} variant of this
pattern. It performs the \emph{reverse conditional} refactoring described 
earlier in both JDT and CDT. The pattern is most visible in the plugin's
swapStatements method, Figure~\ref{fig:swapStatements}, which swaps the 
branches of the \emph{if} statement.

>>>>

% Metadata Information
\acmVolume{0}
\acmNumber{0}
\acmArticle{0}
\articleSeq{0}
\acmYear{2016}
\acmMonth{8}

@@

    \item Functional: \textit{Work with the ASTs in a functional style, where every edit
        returns a new AST.} So, a call to field.removeFromTree() would not
        change the existing AST, but rather it would return a new AST that has
        the given field omitted. When dealing with immutable data structures,
        functional programming may often be the natural solution. However,
        in the case of ASTs, a functional style risks reintroducing many of 
        the problems inherent to mutable trees. File location of nodes may
        be disrupted in the new tree, pointers to nodes in the original tree may
        be rendered 
        irrelevant, semantic information may be invalidated, etc. Those are in
        addition to the difficulties of functional programming, such as memory
        consumption and copying, especially with very large ASTs. This is
        a valid refactoring API design pattern, but any implementation using 
        this variant must carefully consider these problems.

    \item Rewriter: \textit{Provide an API that allows clients to specify what changes
        they want to make in terms of AST nodes.}  This API can then translate these into string edits.  For example, an API might offer a \textit{Rewriter} 
class. The client would inform the \textit{Rewriter} that it wants to 
effectively delete a particular \textit{FieldDeclaration} node from the AST:
\textit{rewriter.delete(field)}.  When it is ready to manipulate the source
code, it delegates this task to the rewriter:
\textit{rewriter.rewrite(sourceCode)}.  This allows multiple edits to be made 
without conflicting with each other. This is illustrated in
Figure~\ref{fig:seq-diagram}.  The Eclipse Java and C/C++ Development
Tools~\cite{cdt-refactoring} take this approach (with slightly different class
and method names).

@@

\section{Example}
We created a plugin for Eclipse that exhibits the \emph{Rewriter} variant of this
pattern\footnote{GitHub Repository: \url{https://github.com/aggrand/reverse-conditional}}. It performs the \emph{reverse conditional} refactoring described 
earlier in both JDT and CDT. The pattern is most visible in the plugin's
swapStatements method, Figure~\ref{fig:swapStatements}, which swaps the 
branches of the \emph{if} statement.
